Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-06-14 paper/ms.tex
Update frozen figs in paper/

\begin{figure*}[h!]%
\centering
\includegraphics[width=18cm]{fig/fig1.pdf}
\caption{\textbf{Synthetic Hubble Space Telescope images of our simulations with
and without the interstellar medium.} The upper panels show an \Nbody{} only
simulation while the lower panels show a simulation which includes the SMUGGLE
model for the interstellar medium. Each panel is $20\,\textrm{kpc}$ to a side.
Columns show different points in time, separated by about $1\,\textrm{Gyr}$.
We can see that in the \Nbody{} run, the bar grows in length and strength. In
the SMUGGLE run, the bar remains at approximately the same length and strength
over the course of the simulation. The \Nbody{} model is identical to the
GALAKOS model, discussed in the text. Details on the production of these
synthetic images is given in the Methods section.}\label{fig:overview}
\end{figure*}

\begin{figure}[h!]%
\centering
\includegraphics[width=9cm]{fig/fig2.pdf}
\caption{\textbf{Properties of the bar in our \boldmath$N$-body and SMUGGLE
simulations of a barred galaxy.} The \textit{upper panel} shows the evolution
of the pattern speed. As expected, the bar in the \Nbody{} run slows down due
to interactions between the bar and the dark matter halo. However, the bar in
the SMUGGLE run does not slow down and instead remains at a constant pattern
speed. The \textit{middle panel} shows the evolution of the bar length. In the
\Nbody{} case, the bar lengthens. This occurs because as the pattern speed
drops, bar-like orbits at larger radii are possible. Stars are captured on
these orbits, lengthening the bar. This process does not occur in the SMUGGLE
cases since the bar pattern speed is not decreasing, and therefore the bar
length remains constant. The \textit{lower panel} shows the torque on the bar
by different components. The solid lines show the torque exerted by the halo
in both the \Nbody{} and SMUGGLE cases. The dashed line shows the torque
exerted by the gas phase in the SMUGGLE run (there is no gas in the \Nbody{}
run). Details on how these properties are calculated is given in the Methods
section.}\label{fig:prop}
\end{figure}

\begin{figure*}[h!]%
\centering
\includegraphics[width=18cm]{fig/fig3.pdf}
\caption{\textbf{The wake excited in the dark matter halo.} The dark matter
halo wake is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
(\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
\textit{middle} and \textit{right panels} show a surface density projection in
the $x$-$y$ plane of the dark matter halo after an axisymmetric average has
been subracted. The solid line indicates the direction of the bar while the
dashed line indicates the direction of the halo wake (both measured by taking
the second Fourier component within a sphere of all material within a radius
of $4\,\textrm{kpc}$). The \textit{left panel} shows the time evolution of
the angle difference between the bar and the halo wake, as measured from the
second Fourier component. After the first Gyr, the angle difference in the
SMUGGLE case is smaller than in the \Nbody{} case by about a factor of two,
reflecting how the dark matter halo in the SMUGGLE case is unable to exert as
negative a torque on the bar as in the \Nbody{} case.}\label{fig:wake}
\end{figure*}

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-snap700.pdf}
\caption{\textbf{Pattern speed evolution with a lower initial pattern speed.} We
tested the evolution of our system when gas is added to the \Nbody{} run at a
later time, but with all other simulation parameters kept the same. The setup
is therefore identical to our previous runs just with a bar that is larger,
stronger, and with a lower pattern speed. We find that the pattern speed
evolution is very similar to our fiducial case, except that the bar retains
its original pattern speed. This indicates a mechanism which keeps the bar at
its formation pattern speed, and that there is not a particular pattern speed
which the system tends to.}
\label{fig:snap700}
\end{figure}

@@

\begin{figure}[t]%
\centering
\includegraphics[width=9cm]{fig/fig-vcirc.pdf}
\caption{\textbf{The circular velocity curve of our initial setups.} This
curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
compared to observational estimates for the Milky
Way.\cite{2019ApJ...871..120E} We see that the circular velocity curve for
both runs is marginally larger than the Milky Way's, but still comparable. The
SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
additional mass in the gas phase.}
\label{fig:vcirc}
\end{figure}

\begin{figure}[h!]%
\centering
\includegraphics[width=9cm]{fig/fig-GFM.pdf}
\caption{\textbf{Pattern speed evolution of a smooth ISM model.} This
evolution is shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE
(orange), and smooth ISM (red) cases. The smooth ISM model is an older model
for the ISM which treats its multiphase nature in a subgrid
fashion.\cite{2003MNRAS.339..289S} This fundamentally differs from the SMUGGLE
model, which explicitly resolve the hot and cold phases of the
ISM.\cite{2019MNRAS.489.4233M} The pattern speed in the smooth ISM case is
broadly similar to the evolution in the SMUGGLE case. This shows that the
stability of the pattern speed is not simply a result of our assumed model for
the ISM.}
\label{fig:GFM}
\end{figure}

@@

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-fgas.pdf}
\caption{\textbf{Pattern speed evolution with varying gas fractions.} We
explored the impact of lowering the initial gas surface density of our
fiducial disk on the evolution of the pattern speed. The surface densities we
tested of $20$, $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to
initial gas fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that
initial surface densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in
bars which remain at a constant pattern speed, while an initial surface
density of $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in
roughly the same manner as the \Nbody{} case. }
\label{fig:fgas}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=9cm]{fig/fig4.pdf}
\caption{\textbf{Average torque exerted by gas on a bar which rotates at a
fixed pattern speed.} Since only gas within the corotation radius is able to
infall and slower bars have larger corotation radii, slower bars experience a
larger net torque than faster bars. The setup of the simulations used here is
identical to the SMUGGLE case discussed earlier and in the Methods section,
except the \Nbody{} disk is rotated as a solid body with a constant angular
velocity.}\label{fig:equil}
\end{figure}

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-surf.pdf}
\caption{\textbf{Evolution of gas and SFR surface density profiles.} The time
evolution of the atomic gas surface density (\textit{upper}), molecular gas
surface density (\textit{middle}) and the star formation rate (SFR) surface
density (\textit{lower}) at various times during our fiducial simulation.
Colored lines indicate the profiles at selected times during the simulation
while the black dashed lines indicate observations for the atomic
gas\cite{2008AA...487..951K} and black dotted lines indicate a model which
allows the CO-to-H$_2$ conversion factor $X_{\textrm{CO}}$ to vary with
metallicity.\cite{2022ApJ...929L..18E} Molecular gas surface densities were
provided separately (N. Evans, private communication). We see that the
molecular gas and SFR surface densities are within an order of magnitude of
the Milky Way's typical values at all times.  We see a sharp decrease in the
gas and SFR surface densities along the extent of the bar from $\sim1$ to
$\sim4\,\textrm{kpc}$, related to the gas inflow in this region.}
\label{fig:surf}
\end{figure}

@@

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-A2.pdf}
\caption{\textbf{Evolution of bar strength.} The bar strength is measured as
the maximum of the second Fourier component divided by the zeroth Fourier
component. We see that in the
\Nbody{} case (blue) the bar strength increases with time, consistent with
previous results showing that the bar strength increases as bars slow down. In
the SMUGGLE case (orange), we see that the bar strength remains roughly
constant, possibly slightly decreasing with time. This is also consistent with
the expected relation between pattern speed and strength since the bar in this
case is not slowing down.}
\label{fig:strength}
\end{figure}

\begin{figure*}[t!]%
\centering
\includegraphics[width=18cm]{fig/fig-bar.pdf}
\caption{\textbf{Disk decomposition into the barred and unbarred disk.} This
procedure is based on ref.\cite{2016MNRAS.463.1952P} The \textit{left panel}
shows a face-on surface density projection through the stellar component of the
\SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The \textit{middle panel} shows the
component of the disk identified as being trapped in the bar while the
\textit{right panel} shows the component of the disk identified as not being
trapped in the bar. The fact that the untrapped stars form a roughly
axisymmetric structure indicates our bar decomposition is sufficiently
accurate.}
\label{fig:decomp}
\end{figure*}

>>>>

\begin{figure*}[h!]%
\centering
\includegraphics[width=18cm]{fig/mockHST.pdf}
\caption{\textbf{Synthetic Hubble Space Telescope images of our simulations with
and without the interstellar medium.} The upper panels show an \Nbody{} only
simulation while the lower panels show a simulation which includes the SMUGGLE
model for the interstellar medium. Each panel is $20\,\textrm{kpc}$ to a side.
Columns show different points in time, separated by about $1\,\textrm{Gyr}$.
We can see that in the \Nbody{} run, the bar grows in length and strength. In
the SMUGGLE run, the bar remains at approximately the same length and strength
over the course of the simulation. The \Nbody{} model is identical to the
GALAKOS model, discussed in the text. Details on the production of these
synthetic images is given in the Methods section.}\label{fig:overview}
\end{figure*}

\begin{figure}[h!]%
\centering
\includegraphics[width=9cm]{fig/ps_Rbar_torque.pdf}
\caption{\textbf{Properties of the bar in our \boldmath$N$-body and SMUGGLE
simulations of a barred galaxy.} The \textit{upper panel} shows the evolution
of the pattern speed. As expected, the bar in the \Nbody{} run slows down due
to interactions between the bar and the dark matter halo. However, the bar in
the SMUGGLE run does not slow down and instead remains at a constant pattern
speed. The \textit{middle panel} shows the evolution of the bar length. In the
\Nbody{} case, the bar lengthens. This occurs because as the pattern speed
drops, bar-like orbits at larger radii are possible. Stars are captured on
these orbits, lengthening the bar. This process does not occur in the SMUGGLE
cases since the bar pattern speed is not decreasing, and therefore the bar
length remains constant. The \textit{lower panel} shows the torque on the bar
by different components. The solid lines show the torque exerted by the halo
in both the \Nbody{} and SMUGGLE cases. The dashed line shows the torque
exerted by the gas phase in the SMUGGLE run (there is no gas in the \Nbody{}
run). Details on how these properties are calculated is given in the Methods
section.}\label{fig:prop}
\end{figure}

\begin{figure*}[h!]%
\centering
\includegraphics[width=18cm]{fig/halo_wake.pdf}
\caption{\textbf{The wake excited in the dark matter halo.} The dark matter
halo wake is shown in the \Nbody{} case (\textit{middle panel}) and SMUGGLE case
(\textit{right panel}) after $2.6\,\textrm{Gyr}$ of evolution. The
\textit{middle} and \textit{right panels} show a surface density projection in
the $x$-$y$ plane of the dark matter halo after an axisymmetric average has
been subracted. The solid line indicates the direction of the bar while the
dashed line indicates the direction of the halo wake (both measured by taking
the second Fourier component within a sphere of all material within a radius
of $4\,\textrm{kpc}$). The \textit{left panel} shows the time evolution of
the angle difference between the bar and the halo wake, as measured from the
second Fourier component. After the first Gyr, the angle difference in the
SMUGGLE case is smaller than in the \Nbody{} case by about a factor of two,
reflecting how the dark matter halo in the SMUGGLE case is unable to exert as
negative a torque on the bar as in the \Nbody{} case.}\label{fig:wake}
\end{figure*}

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/ps_late_start.pdf}
\caption{\textbf{Pattern speed evolution with a lower initial pattern speed.} We
tested the evolution of our system when gas is added to the \Nbody{} run at a
later time, but with all other simulation parameters kept the same. The setup
is therefore identical to our previous runs just with a bar that is larger,
stronger, and with a lower pattern speed. We find that the pattern speed
evolution is very similar to our fiducial case, except that the bar retains
its original pattern speed. This indicates a mechanism which keeps the bar at
its formation pattern speed, and that there is not a particular pattern speed
which the system tends to.}
\label{fig:snap700}
\end{figure}

@@

\begin{figure}[t]%
\centering
\includegraphics[width=9cm]{fig/vcirc.pdf}
\caption{\textbf{The circular velocity curve of our initial setups.} This
curve is shown for the \Nbody{} run (blue) and the SMUGGLE run (orange)
compared to observational estimates for the Milky
Way.\cite{2019ApJ...871..120E} We see that the circular velocity curve for
both runs is marginally larger than the Milky Way's, but still comparable. The
SMUGGLE circular velocity curve is larger than the \Nbody{} curve due to the
additional mass in the gas phase.}
\label{fig:vcirc}
\end{figure}

\begin{figure}[h!]%
\centering
\includegraphics[width=9cm]{fig/ps_GFM.pdf}
\caption{\textbf{Pattern speed evolution of a smooth ISM model.} This
evolution is shown for the fiducial disk in the \Nbody{} (blue), SMUGGLE
(orange), and smooth ISM (red) cases. The smooth ISM model is an older model
for the ISM which treats its multiphase nature in a subgrid
fashion.\cite{2003MNRAS.339..289S} This fundamentally differs from the SMUGGLE
model, which explicitly resolve the hot and cold phases of the
ISM.\cite{2019MNRAS.489.4233M} The pattern speed in the smooth ISM case is
broadly similar to the evolution in the SMUGGLE case. This shows that the
stability of the pattern speed is not simply a result of our assumed model for
the ISM.}
\label{fig:GFM}
\end{figure}

@@

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/ps_fgas.pdf}
\caption{\textbf{Pattern speed evolution with varying gas fractions.} We
explored the impact of lowering the initial gas surface density of our
fiducial disk on the evolution of the pattern speed. The surface densities we
tested of $20$, $15$, $10$, and $5\,\Msun/\textrm{kpc}^2$ correspond to
initial gas fractions of $16\%$, $10\%$, $7\%$, and $4\%$. We find that
initial surface densities $20$, $15$, and $10\,\Msun/\textrm{kpc}^2$ result in
bars which remain at a constant pattern speed, while an initial surface
density of $5\,\Msun/\textrm{kpc}^2$ results in a bar which slows down in
roughly the same manner as the \Nbody{} case. }
\label{fig:fgas}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=9cm]{fig/torque_ps.pdf}
\caption{\textbf{Average torque exerted by gas on a bar which rotates at a
fixed pattern speed.} Since only gas within the corotation radius is able to
infall and slower bars have larger corotation radii, slower bars experience a
larger net torque than faster bars. The setup of the simulations used here is
identical to the SMUGGLE case discussed earlier and in the Methods section,
except the \Nbody{} disk is rotated as a solid body with a constant angular
velocity.}\label{fig:equil}
\end{figure}

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/surf_dens.pdf}
\caption{\textbf{Evolution of gas and SFR surface density profiles.} The time
evolution of the atomic gas surface density (\textit{upper}), molecular gas
surface density (\textit{middle}) and the star formation rate (SFR) surface
density (\textit{lower}) at various times during our fiducial simulation.
Colored lines indicate the profiles at selected times during the simulation
while the black dashed lines indicate observations for the atomic
gas\cite{2008AA...487..951K} and black dotted lines indicate a model which
allows the CO-to-H$_2$ conversion factor $X_{\textrm{CO}}$ to vary with
metallicity.\cite{2022ApJ...929L..18E} Molecular gas surface densities were
provided separately (N. Evans, private communication). We see that the
molecular gas and SFR surface densities are within an order of magnitude of
the Milky Way's typical values at all times.  We see a sharp decrease in the
gas and SFR surface densities along the extent of the bar from $\sim1$ to
$\sim4\,\textrm{kpc}$, related to the gas inflow in this region.}
\label{fig:surf}
\end{figure}

@@

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/A2_over_A0.pdf}
\caption{\textbf{Evolution of bar strength.} The bar strength is measured as
the maximum of the second Fourier component divided by the zeroth Fourier
component. We see that in the
\Nbody{} case (blue) the bar strength increases with time, consistent with
previous results showing that the bar strength increases as bars slow down. In
the SMUGGLE case (orange), we see that the bar strength remains roughly
constant, possibly slightly decreasing with time. This is also consistent with
the expected relation between pattern speed and strength since the bar in this
case is not slowing down.}
\label{fig:strength}
\end{figure}

\begin{figure*}[t!]%
\centering
\includegraphics[width=18cm]{fig/bar_decomp.pdf}
\caption{\textbf{Disk decomposition into the barred and unbarred disk.} This
procedure is based on ref.\cite{2016MNRAS.463.1952P} The \textit{left panel}
shows a face-on surface density projection through the stellar component of the
\SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The \textit{middle panel} shows the
component of the disk identified as being trapped in the bar while the
\textit{right panel} shows the component of the disk identified as not being
trapped in the bar. The fact that the untrapped stars form a roughly
axisymmetric structure indicates our bar decomposition is sufficiently
accurate.}
\label{fig:decomp}
\end{figure*}
